Use of management strategy evaluation to inform fisheries management in a changing climate

### 2019-09-03 Poster_ICES.Rmd
Edit header font; edit caption font size; Change name of reviewed to random sample; edit colors; edit text based on feedback

---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0
#ESSENTIALS
title: Use of management strategy evaluation to inform fisheries management in a changing climate
author:
  - name: '**Jonathan W. Cummings**'
    affil: 1
    main: true
    orcid: '0000-0001-8028-5787'
    twitter: jonwcummings 
    email: 'jcummings@umassd.edu'
  - name: Amanda Hart
    affil: 1
    main: true
    orcid: '0000-0002-6449-9328'
    email: 'ahart2@umassd.edu'
  - name: Gavin Fay
    main: true
    affil: 1
    email: 'gfay@umassd.edu'
affiliation:
  - num: 1
    address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
main_findings: "Explicit documentation of management strategy evaluation processes is limited. <br> <br> Climate change MSEs have been less explicit, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: MSEreview.bib
link-citations: true
nocite: '@*'
title_textsize: "50pt"
main_textsize: "90pt"
column_numbers: 3
body_textsize: "29.5pt"
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html")
##### load libraries #####
library(tidyverse) # upgrade to base R
library(readxl) # read in excel files
library(shiny) # Shiny App
library(shinythemes) # web themes for Shiny
library(DT) # Data Table
library(purrr) # for loop alternative
library(ggmap) # map plotting
library(mapdata) # basemap generation
library(odbc) # database connection 
library(RMySQL) # MySQL scripting in R
library(DBI) # database interface
library(kableExtra) # extra formating of the tables

@@

# Where MSEs have occured
map.data<-rbind(data_pub,data_CC) %>%
  select(map.col) %>%
  mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
  mutate(Drivers=replace_na(Drivers, "Other"))

# label MSEs articles by analysis for ploting
freq.data_pub<-mutate(freq.data_pub,Analysis="Reviewed")
freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
freq.data<-rbind(freq.data_pub,freq.data_CC)
part.data_pub<-mutate(part.data_pub,Analysis="Reviewed")
part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
part.data<-rbind(part.data_pub,part.data_CC) %>% 
  arrange(Stage, Percent) %>% 
  mutate(order = row_number())
part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,
                   16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,
                   32,33,34,35,36,37,38,39,39)
drive.data_pub<-mutate(drive.data_pub,Analysis="Reviewed") %>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,Driver,Percent,Frequency)
drive.data_CC<-mutate(drive.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,Driver,Percent,Frequency)
drive.data<-rbind(drive.data_pub,drive.data_CC) %>% 
  arrange(Percent) %>%
  mutate(order = row_number())
drive.data$order<-c(6,5,4,3,2,1,8,7,12,9,3,15,14,13,12,11,10,14,16,15,
                   16,6)
# Wrangle objective category data by reviewed, climate change, and combined
objcat.data_pub<-mutate(objcat.data_pub,Analysis="Reviewed") %>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,'Objective Category',Percent,Frequency)
objcat.data_CC<-mutate(objcat.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,'Objective Category',Percent,Frequency)
objcat.data<-rbind(objcat.data_pub,objcat.data_CC) %>% 
  rename(OC='Objective Category') %>% 
  arrange(OC) %>% 
  rename('Objective Category'=OC)

# Wrangle objective data by reviewed, climate change, and combined
obj.dataTable_pub<-mutate(obj.dataTable_pub,Analysis="Reviewed") %>% 
  mutate(Percent=round(Percent,0))
obj.dataTable_CC<-mutate(obj.dataTable_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0))
obj.dataTable<-rbind(obj.dataTable_pub,obj.dataTable_CC) %>% 
  filter(Objective=="Category") %>%
  group_by(Analysis) %>% 
  summarise('AVG'=sum(Number)) %>% 
  mutate(count=c(11,30)) %>% 
  mutate('Per MSE'=round(AVG/count,1)) %>% 
  mutate('Type'=c("Objectives","Objectives"))

altcat.data_pub<-mutate(altcat.data_pub,Analysis="Reviewed") %>% 
  mutate(Percent=round(Percent,0))
altcat.data_CC<-mutate(altcat.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0))
altcat.data<-rbind(altcat.data_pub,altcat.data_CC) %>%
  group_by(Analysis) %>% 
  summarise('AVG'=sum(Number)) %>% 
  mutate(count=c(11,30)) %>% 
  mutate('Per MSE'=round(AVG/count,1)) %>% 
  mutate('Type'=c("Alternatives","Alternatives"))

@@

# plot MSEs on map
MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
  geom_point(size=2) + scale_color_manual(values=c("#FF0033","#0000FF")) +  theme_void() +
  theme(legend.position = c(0.15, 0.25))

# plot explicit documentation of steps or components of MSE processes
# Reviewed MSEs
Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
  geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
  geom_vline(xintercept=1.5,linetype="dashed")+
  coord_flip()+scale_fill_manual(values=c("#FF0033","#0000FF")) +
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
  scale_x_discrete(
    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process"),
    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process")) +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.875,0.085),legend.title = element_blank())

Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
  facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
  scale_fill_manual(values=c("#FF0033","#0000FF")) +
  scale_x_continuous(breaks = part.data$order,
                     labels = part.data$Participants)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
  ylab("Percent")+xlab(NULL)+coord_flip() +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.875,0.085),legend.title = element_blank())

Driver.plot<-ggplot(drive.data,aes(x=order,y=Percent,fill=Analysis)) +
  geom_col(position="dodge") +
  scale_fill_manual(values=c("#FF0033","#0000FF")) +
  scale_x_continuous(breaks = drive.data$order,
                     labels = drive.data$Driver)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+
  ylab("Percent")+xlab(NULL)+coord_flip() +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.85,0.085),legend.title = element_blank())
```

# Introduction
For most of its history fisheries management has focused on fishing impacts with ecosystem status viewed as a background constant. The increasing rate and degree of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.

Management strategy evaluation (MSE) is commonly considered the gold standard of fisheries management approaches, and is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt]. Thus, MSE is a compelling tool to address the impact of climate change on fisheries management.

Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fisheries management that accounts for a changing environment.

Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives. Using the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?

@@

## Objectives
We review published MSEs against the steps of the SDM framework to:

* Determine the rate of documentation and use of steps in the SDM process.
* Identify participants in MSE processes.
* Highlight unique aspects of MSEs that model climate change as a driver.
* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
* Provide insights about MSE processes and documentation that can further learning to improve future MSEs

# Methods
#### Published MSE article selection
We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019. This search returned 253 results.

We took a random sample of these 253 articles and reviewed 30 that document a MSE.  Not all articles in the potential pool of 253 articles were MSEs due to articles that are reviews, meta-analyses, etc., or simply cite other MSE articles.  Of the 253 articles, ### documented a MSE, and 11 included climate change as a driver of fishery status.

| MSE type | Count |
|:--:|:--:|
| Published | ### |
| Reviewed Sample | 30 |
| Climate Change | 11 |

</br>
</br>

# Results
Apart from one each in Europe and the Western Pacific, to date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).

```{r, mapfigure, fig.cap='Map of MSE study locations.', out.width="100%"}
MSE.map
```

The vast majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).

```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results. In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=6, out.width="100%"}
text1<-expression(paste("Inexplicit \nor Absent"))
text2<-paste("Absent")
text3<-paste("Unknown")
Freq.plot + annotate("text", x = 5, y = 55, label = text1, size=25) + 
  annotate("text", x = 2.6, y = 50, label = text2, size=20) + 
  annotate("text", x = 1, y = 50, label = text3, size=12)
```

The primary participants and leaders of the MSE processes were scientists, while some processes were lead by governments and management agencies (Figure \@ref(fig:participants)).  Where explicit documentation occurred objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in the inexplicit cases.

```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
Part.plot
```

General uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the reviewed MSEs (Figure \@ref(fig:driversPlot)).

```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"}
Driver.plot
```

A greater number of objectives were considered per MSE in the random sample of reviewed MSEs than in the climate change MSEs (Table \@ref(tab:PerMSE)).  Climate change MSEs were less likely to consider economic and social objectives than the MSEs sample for review (Table \@ref(tab:objectiveCategories)).

```{r, PerMSE}
color_Pub <- which(per.MSE$Analysis=='Reviewed')
color_CC <- which(per.MSE$Analysis=='Climate Change')

per.MSE %>%
  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%  
  row_spec(color_CC, color = "#FF0033", background = "white") %>%
  row_spec(color_Pub, color = "#0000FF", background = "white")
```

```{r, objectiveCategories}
color_Pub <- which(objcat.data$Analysis=='Reviewed')
color_CC <- which(objcat.data$Analysis=='Climate Change')

objcat.data %>%
  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
  row_spec(color_CC, color = "#FF0033", background = "white") %>%
  row_spec(color_Pub, color = "#0000FF", background = "white")
```

# Conclusions
The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. While documentation about the modeling conducted in a MSE was explicit, documentation about the process used to make decision about the MSE model were limited. Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs. Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.  Conservation, Economic, and Yield focused objectives were commonly considered, while broader social objectives were rarely considered.

To date climate change has not been considered in many MSEs.  MSEs modeling climate change have been less explicitly documented, more likely to be scientist driven, and included fewer system drivers and fewer objectives, with less consideration of economic or social objectives.

# More Details
Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>) and keep an eye out for the forthcoming publication.

# References

>>>>

---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0
#ESSENTIALS
title: Use of management strategy evaluation to inform fisheries management in a changing climate
author:
  - name: '**Jonathan W. Cummings**'
    affil: 1
    main: true
    orcid: '0000-0001-8028-5787'
    twitter: jonwcummings 
    email: 'jcummings@umassd.edu'
  - name: Amanda Hart
    affil: 1
    main: true
    orcid: '0000-0002-6449-9328'
    email: 'ahart2@umassd.edu'
  - name: Gavin Fay
    main: true
    affil: 1
    email: 'gfay@umassd.edu'
affiliation:
  - num: 1
    address: School for Marine Science and Technology, University of Massachusetts - Dartmouth 836 South Rodney French Blvd,New Bedford, MA, USA 02744
main_findings: "Explicit documentation of management strategy evaluation processes is limited. <br> <br> Climate change MSEs have been less explicit, more scientist driven, included fewer objectives -- with less consideration of economic or social objectives -- and fewer drivers."
output: 
  posterdown::posterdown_betterport:
    self_contained: false
    pandoc_args: --mathjax
    number_sections: false
bibliography: MSEreview.bib
link-citations: true
nocite: '@*'
title_textsize: "50pt"
main_textsize: "90pt"
column_numbers: 3
body_textsize: "29.5pt"
main_fontfamily: "Rasa"
caption_textsize: "24pt"
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html")
##### load libraries #####
library(tidyverse) # upgrade to base R
library(readxl) # read in excel files
library(shiny) # Shiny App
library(shinythemes) # web themes for Shiny
library(DT) # Data Table
library(purrr) # for loop alternative
library(ggmap) # map plotting
library(mapdata) # basemap generation
library(odbc) # database connection 
library(RMySQL) # MySQL scripting in R
library(DBI) # database interface
library(kableExtra) # extra formating of the tables

@@

# Where MSEs have occured
map.data<-rbind(data_pub,data_CC) %>%
  select(map.col) %>%
  mutate(Drivers=str_extract(Drivers, "Climate Change")) %>%
  mutate(Drivers=replace_na(Drivers, "Random Sample"))

# label MSEs articles by analysis for ploting
freq.data_pub<-mutate(freq.data_pub,Analysis="Random Sample")
freq.data_CC<-mutate(freq.data_CC,Analysis="Climate Change")
freq.data<-rbind(freq.data_pub,freq.data_CC)
part.data_pub<-mutate(part.data_pub,Analysis="Random Sample")
part.data_CC<-mutate(part.data_CC,Analysis="Climate Change")
part.data<-rbind(part.data_pub,part.data_CC) %>% 
  arrange(Stage, Percent) %>% 
  mutate(order = row_number())
part.data$order<-c(1,2,3,3,4,4,5,6,7,8,6,9,10,8,9,11,12,10,13,12,13,14,14,15,16,17,18,19,20,
                   16,17,18,20,21,22,23,24,25,21,23,26,25,26,27,27,28,29,30,31,29,32,30,33,
                   32,33,34,35,36,37,38,39,39)
drive.data_pub<-mutate(drive.data_pub,Analysis="Random Sample") %>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,Driver,Percent,Frequency)
drive.data_CC<-mutate(drive.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,Driver,Percent,Frequency)
drive.data<-rbind(drive.data_pub,drive.data_CC) %>% 
  arrange(Percent) %>%
  mutate(order = row_number())
drive.data$order<-c(6,5,4,3,2,1,8,7,12,9,3,15,14,13,12,11,10,14,16,15,
                   16,6)
# Wrangle objective category data by Random Sample, climate change, and combined
objcat.data_pub<-mutate(objcat.data_pub,Analysis="Random Sample") %>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,'Objective Category',Percent,Frequency)
objcat.data_CC<-mutate(objcat.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0)) %>% 
  select(Analysis,'Objective Category',Percent,Frequency)
objcat.data<-rbind(objcat.data_pub,objcat.data_CC) %>% 
  rename(OC='Objective Category') %>% 
  arrange(OC) %>% 
  rename('Objective Category'=OC)

# Wrangle objective data by Random Sample, climate change, and combined
obj.dataTable_pub<-mutate(obj.dataTable_pub,Analysis="Random Sample") %>% 
  mutate(Percent=round(Percent,0))
obj.dataTable_CC<-mutate(obj.dataTable_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0))
obj.dataTable<-rbind(obj.dataTable_pub,obj.dataTable_CC) %>% 
  filter(Objective=="Category") %>%
  group_by(Analysis) %>% 
  summarise('AVG'=sum(Number)) %>% 
  mutate(count=c(11,30)) %>% 
  mutate('Per MSE'=round(AVG/count,1)) %>% 
  mutate('Type'=c("Objectives","Objectives"))

altcat.data_pub<-mutate(altcat.data_pub,Analysis="Random Sample") %>% 
  mutate(Percent=round(Percent,0))
altcat.data_CC<-mutate(altcat.data_CC,Analysis="Climate Change")%>% 
  mutate(Percent=round(Percent,0))
altcat.data<-rbind(altcat.data_pub,altcat.data_CC) %>%
  group_by(Analysis) %>% 
  summarise('AVG'=sum(Number)) %>% 
  mutate(count=c(11,30)) %>% 
  mutate('Per MSE'=round(AVG/count,1)) %>% 
  mutate('Type'=c("Alternatives","Alternatives"))

@@

# plot MSEs on map
MSE.map<-ggplot(data=map.data,aes(x=Longitude, y=Latitude,color=Drivers)) + world +
  geom_point(size=2.5) + scale_color_manual(values=c("#66CCCC","#006666")) +  theme_void() +
  theme(legend.position = c(0.15, 0.25))

# plot explicit documentation of steps or components of MSE processes
# Random Sample MSEs
Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
  geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
  geom_vline(xintercept=1.5,linetype="dashed")+
  coord_flip()+scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
  scale_x_discrete(
    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process"),
    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process")) +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.85,0.07),legend.title = element_blank())

Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
  facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
  scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_x_continuous(breaks = part.data$order,
                     labels = part.data$Participants)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
  ylab("Percent")+xlab(NULL)+coord_flip() +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.86,0.075),legend.title = element_blank())

Driver.plot<-ggplot(drive.data,aes(x=order,y=Percent,fill=Analysis)) +
  geom_col(position="dodge") +
  scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_x_continuous(breaks = drive.data$order,
                     labels = drive.data$Driver)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+
  ylab("Percent")+xlab(NULL)+coord_flip() +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.825,0.075),legend.title = element_blank())
```

###
Fisheries management has mostly focused on fishing impacts with ecosystem status viewed as a background constant.

The increasing rate of climate change is changing this dynamic, bringing ecosystem status to the forefront of fisheries management.

Management strategy evaluation (MSE) is "widely considered to be the most appropriate way to evaluate the trade‐offs achieved by alternative management strategies and to assess the consequences of uncertainty for achieving management goals” [@R-Punt]. Thus, MSE is a compelling tool to assess climate change impacts and test climate-ready options for fisheries management decisions.

Adaptive management arose to address uncertainties and accelerate progress towards meeting management objectives. We used the structured decision making (SDM) process [@R-SDM] -- the decision making framework in which adaptive management occurs -- as our framework, how do published MSE projects utilize standard SDM components and support learning within the MSE practitioner community?

<center>
![SDM process](SDM process.png)

@@

### Objectives
We review published MSEs against the steps of the SDM framework to:

* Determine the rate of documentation, and use of steps, in the MSE process:  
* Identify participants in MSE processes.
* Highlight unique aspects of MSEs that model climate change as a driver.
* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
* Provide insights about MSE processes and documentation that can further learning to improve future MSEs

### MSE Literature Review Methods
We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019. This search returned 253 results.  We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample. Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.

| MSE type | Count |
|:--:|:--:|
| Published (estimated) | 110-160 |
| Random Sample | 30 |
| Climate Change | 11 |  
<br>
<br>

### Where have climate change MSEs occurred?
To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)).

```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"}
MSE.map +
  theme(text = element_text(size=14))
```

### Are structured decision making steps explicit in MSEs?
The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Figure \@ref(fig:explicit)).

```{r explicit, fig.cap='Percentage of MSEs that included explicit documentation (colored bars) of SDM process steps --- Roles and meetings --- and adoption of results. In the remaining MSEs was documentation was inexplicit, or absent, and whether the results were adopted was unknown.', fig.height=5.5, out.width="100%"}
text1<-expression(paste("Inexplicit \nor Absent"))
text2<-paste("Absent")
text3<-paste("Unknown")
Freq.plot + annotate("text", x = 8.75, y = 50, label = 'underline("Else")', size=20, parse=TRUE) +
  annotate("text", x = 5, y = 52, label = text1, size=20) + 
  annotate("text", x = 2.6, y = 50, label = text2, size=20) + 
  annotate("text", x = 1, y = 50, label = text3, size=12) +
  theme(text = element_text(size=14))
```

### Who is involved in MSEs?
The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).  Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.

```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
Part.plot +
  theme(text = element_text(size=14))
```

### What questions and management objectives do MSEs address?
Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)).

```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"}
Driver.plot +
  theme(text = element_text(size=14))
```

More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).  Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)).

```{r, PerMSE}
color_Pub <- which(per.MSE$Analysis=='Random Sample')
color_CC <- which(per.MSE$Analysis=='Climate Change')

per.MSE %>%
  kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%  
  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
  row_spec(color_Pub, color = "#006666", background = "white")
```

```{r, objectiveCategories}
color_Pub <- which(objcat.data$Analysis=='Random Sample')
color_CC <- which(objcat.data$Analysis=='Climate Change')

objcat.data %>%
  kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>%
  row_spec(color_CC, color = "#66CCCC", background = "white") %>%
  row_spec(color_Pub, color = "#006666", background = "white")
```

### Main Conclusions
The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**. Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs. Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.  Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.

Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.

### Explore our review on the web!
Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).

### References
